---
title: 'Gut Diversity'
author: "X"
date: '2023-05-30'
output:
  html_document:
    number_sections: yes
    toc: yes
    theme: united
    highlight: tango
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE, 
                      fig.align = "center", 
                      fig.width = 8,
                      fig.height = 6, 
                      message = FALSE, 
                      warning = FALSE)
```

# Libraries
```{r libraries}
library(metabaR)
library(ggplot2)
library(vegan)
library(lme4)
library(lmerTest)
library(emmeans)
library(dplyr)
library(reshape)
library(reshape2)
library(purrr)
```

# Read the data and load functions
```{r, echo=TRUE}
setwd("~/data")

all_data <- readRDS("FrommMS1_cleandata.rds")

data_lizards <- all_data$samples

setwd("~/R")
source('0_Fromm_sources_functions.R', chdir = TRUE)
```

# Rarefaction curves 

```{r}
# run rarefaction (use boot = 20 to limit computation time)
all_data.raref <- hill_rarefaction(all_data, nboot = 20, nsteps = 10)

gghill_rarefaction(
         all_data.raref,
         group = mblist_Emma$samples$Climate_TM1[match(all_data$pcrs$sample_id,
                                               rownames(all_data$samples))]) +
  labs(color = "Traitement", x = "Reads") + 
  scale_fill_manual(values = c("#006699", "#CC0000")) +
  scale_color_manual(values = c("#006699", "#CC0000"), labels=c("Present-day", "Warm")) +
  theme_fig + 
  theme(legend.position = "bottom",
        legend.title = element_text(size = 0), 
        legend.text = element_text(size=14), 
        axis.text.x = element_text(size=12))
```

# 1 : Alpha diversity [exp(H)]

## Plot

```{r, echo=FALSE, fig.align='center'}
graphShannon_global <- ggplot(data_lizards, mapping = aes(x = Climate_TM1, y = q1, fill=Climate_TM1)) +
  geom_boxplot(outlier.shape = NA, alpha=0.5) +
  geom_point(aes(color=Climate_TM1), 
             position = position_jitterdodge(), 
             alpha=0.35)+
  facet_grid(cols =  vars(Year), rows = vars(Connectivity_TM1)) + 
  labs(y="exp(H)", x = "") +
  scale_x_discrete(labels = c("Present-day", "Warm")) +
  scale_fill_manual(name = "Climatic treatment", labels = c("Present-day", "Warm"), values=c("#006699", "#CC0000")) + 
  scale_color_manual(values=c("#006699", "#CC0000")) + 
  theme_fig 

graphShannon_global
```

```{r, eval=FALSE}
#to save
setwd("~/Plots")
ggsave("graphShannon_global.pdf", graphShannon_global, width = 8, height = 5)
```

## Initial conditions (year0)

```{r, echo=FALSE}
data_2015 <- subset(data_lizards, Year=="2015") #Data frame including only lizards year0 (=2015)
data_2015$Mesocosm_TM1 <- as.factor(data_2015$Mesocosm_TM1)
```

### Model year0
```{r, message=FALSE, echo=TRUE}
mod_2015 <- lmer("q1 ~ Climate_TM1*Connectivity_TM1 + Age_class + Sex + SVL + (1|Mesocosm_TM1)", data = data_2015)

dredge_function(mod_2015)
relative_importance(mod_2015)

r.squaredGLMM(mod_2015)

#pairwise t-test
means_2015 = emmeans(mod_2015,~Climate_TM1*Connectivity_TM1) 
contrast(means_2015, method = "pairwise", adjust = "bonferroni", by=c("Connectivity_TM1"))

#validity requirements 
plot(mod_2015)
qqnorm(residuals(mod_2015))
```

## Full model : isolated and connected mesocosms

```{r, echo=FALSE}
#data
data_wo_2015 = data_lizards  %>%
  filter(Year != "2015")

data_wo_2015$Mesocosms_TM1<- as.factor(data_wo_2015$Mesocosm_TM1)
```

### Plot
```{r}
graphShannon <- ggplot(data_wo_2015, mapping = aes(x = Climate_TM1, y = q1, fill=Climate_TM1)) +
  geom_boxplot(outlier.shape = NA, alpha=0.5) +
  geom_point(aes(color=Climate_TM1), 
             position = position_jitterdodge(), 
             alpha=0.35)+
  facet_grid(cols =  vars(Year), rows = vars(Connectivity_TM1)) + 
  labs(y="exp(H)", x = "") +
  scale_x_discrete(labels = c("Present-day", "Warm")) +
  scale_fill_manual(name = "Climatic treatment", labels = c("Present-day", "Warm"), values=c("#006699", "#CC0000")) + 
  scale_color_manual(values=c("#006699", "#CC0000")) + 
  theme_fig 

graphShannon
```

```{r, eval=FALSE}
#Save plot
setwd("~/Plots")
ggsave("graphShannon.pdf", graphShannon, width = 8, height = 5)
```

### Run model

We recode and center the year factor (Schielzeth et al. 2010). 
```{r, echo=FALSE}
data_wo_2015$year2 <- ifelse(data_wo_2015$Year =="2017", 
                            1, 
                            0)


data_wo_2015$year3 <- ifelse(data_wo_2015$Year =="2018", 
                            1, 
                            0)


data_wo_2015$year2_c <- data_wo_2015$year2 - mean(data_wo_2015$year2)
data_wo_2015$year3_c <- data_wo_2015$year3 - mean(data_wo_2015$year3)
```

```{r, echo=TRUE}
mod_wo_2015 <- lmer("q1 ~ year2_c*Climate_TM1*Connectivity_TM1 + year3_c*Climate_TM1*Connectivity_TM1 + Age_class + Sex + SVL + (1|Mesocosm_TM1) + (1|Code)", data = data_wo_2015)

dredge_function(mod_wo_2015)
relative_importance(mod_wo_2015)

r.squaredGLMM(mod_wo_2015)
```

## Model in isolated mesocosms

```{r, echo=FALSE}
#data
data_wo2015_isolated = data_wo_2015 %>%
  filter(Connectivity_TM1 != "Connected")

summary(data_wo2015_isolated$Year)
data_wo2015_isolated$Year <- as.factor(data_wo2015_isolated$Year)

#center years
data_wo2015_isolated$year2 <- ifelse(data_wo2015_isolated$Year =="2017", 
                            1, 
                            0)


data_wo2015_isolated$year3 <- ifelse(data_wo2015_isolated$Year =="2018", 
                            1, 
                            0)


data_wo2015_isolated$year2_c <- data_wo2015_isolated$year2 - mean(data_wo2015_isolated$year2)
data_wo2015_isolated$year3_c <- data_wo2015_isolated$year3 - mean(data_wo2015_isolated$year3)
```

```{r, echo=TRUE}
mod_wo2015_isolated <- lmer("q1 ~  year2_c*Climate_TM1 + year3_c*Climate_TM1 + Age_class + Sex + SVL + (1|Mesocosm_TM1) + (1|Code)", data = data_wo2015_isolated)

dredge_function(mod_wo2015_isolated)
relative_importance(mod_wo2015_isolated)

r.squaredGLMM(mod_wo2015_isolated)
```

## Model in connected mesocosms

```{r, echo=FALSE}
#data
data_connected_wo2015 <- subset(data_wo_2015, Connectivity_TM1 == "Connected")

#center years
data_connected_wo2015$year2 <- ifelse(data_connected_wo2015$Year =="2017", 
                            1, 
                            0)


data_connected_wo2015$year3 <- ifelse(data_connected_wo2015$Year =="2018", 
                            1, 
                            0)


data_connected_wo2015$year2_c <- data_connected_wo2015$year2 - mean(data_connected_wo2015$year2)
data_connected_wo2015$year3_c <- data_connected_wo2015$year3 - mean(data_connected_wo2015$year3)
```

```{r}
mod_connected_wo2015 <- lmer("q1 ~year2_c*Climate_TM1 + year3_c*Climate_TM1 + Age_class + Sex + SVL + (1|Mesocosm_TM1) + (1|Code)", data = data_connected_wo2015)

dredge_function(mod_connected_wo2015)
relative_importance(mod_connected_wo2015)

r.squaredGLMM(mod_connected_wo2015)
```


# 2 : Alpha diversity within major phyla

We run repeated-measure mixed models. 

## Data
```{r}
valeurs <- melt(data_lizards, id= c("Climate_TM1", "Code", "Year", "Connectivity_TM1", "Mesocosm_TM1", "Age_class", "Sex", "SVL"), measure.vars = c("q1_firmi", "q1_bactero", "q1_actino", "q1_fuso", "q1_alphapro", "q1_deltapro", "q1_gammapro"))

valeurs$phylum_firmi <- ifelse(valeurs$variable=="q1_firmi", 
                            "Firmicutes", 
                            NA)
valeurs$phylum_bactero <- ifelse(valeurs$variable=="q1_bactero", 
                            "Bacteroidetes", 
                            valeurs$phylum_firmi)
valeurs$phylum_actino <- ifelse(valeurs$variable=="q1_actino", 
                            "Actinobacteria", 
                            valeurs$phylum_bactero)
valeurs$phylum_fuso <- ifelse(valeurs$variable=="q1_fuso", 
                            "Fusobacteria", 
                            valeurs$phylum_actino)
valeurs$phylum_alphapro <- ifelse(valeurs$variable=="q1_alphapro", 
                            "Alphaproteobacteria", 
                            valeurs$phylum_fuso)
valeurs$phylum_deltapro <- ifelse(valeurs$variable=="q1_deltapro", 
                            "Deltaproteobacteria", 
                            valeurs$phylum_alphapro)
valeurs$phylum <- ifelse(valeurs$variable=="q1_gammapro", 
                            "Gammaproteobacteria", 
                            valeurs$phylum_deltapro)

valeurs_div <- subset(valeurs, select = c("Climate_TM1", "Connectivity_TM1", "Code", "Year", "variable", "value", "phylum", "Mesocosm_TM1", "Age_class", "Sex", "SVL"))

# Create a list
valeurs_div$Year <- as.character(valeurs_div$Year)
list_valeursPhylum <- split(valeurs_div, valeurs_div$phylum)
names(list_valeursPhylum) <- c("A. Actinobacteria", "B. Alphaproteobacteria", "C. Bacteroidetes", "D. Deltaproteobacteria", "E. Firmicutes", "Fusobacteria", "F. Gammaproteobacteria")
```

## Plots by phylum 

```{r}
#Define plot
Plot_phylum <- function(x) {ggplot(list_valeursPhylum[[x]], mapping = aes(x = Year, y = value, fill = Climate_TM1)) + 
    geom_boxplot(outlier.shape = NA, alpha=0.5) + 
    geom_point(aes(color=Climate_TM1), 
             position = position_jitterdodge(), 
             alpha=0.35)+
  facet_grid(cols = vars(Connectivity_TM1)) +
  labs(x="", y="exp(H)") + 
  ggtitle(names(list_valeursPhylum)[x]) + 
  scale_fill_manual(values=c("#006699", "#CC0000"), name="Climatic condition", labels=c("Present", "Warm")) +
  scale_color_manual(values=c("#006699", "#CC0000")) +
  theme_fig_by_phylum}

Plots_out <- lapply(seq_along(list_valeursPhylum), Plot_phylum)
names(Plots_out) <- c("Div_Actino", "Div_Alphapro", "Div_Bactero", "Div_Deltapro", "Div_Firmi", "Div_Fuso", "Div_Gamma")
```

```{r, eval=FALSE}
#save plots
setwd("~/Plots")
iwalk(Plots_out, ~ ggsave(glue::glue("{.y}.png"), .x, device = "png"))
```

## Models 
```{r}
#data without year0
valeurs_wo2015 <- subset(valeurs_div, Year != "2015")
```

### Isolated mesocosms
```{r}
valeurs_isolated <- subset(valeurs_wo2015, Connectivity_TM1 == "Isolated")

valeurs_isolated$Climate_TM1 <- relevel(as.factor(valeurs_isolated$Climate_TM1), ref = "T38")

mod_phylum_isolated <- lmer(value ~ Climate_TM1*phylum*Year + Age_class + Sex + SVL + (1|Code) + (1|Mesocosm_TM1), data = valeurs_isolated)

means_phylum_isolated  <- emmeans(mod_phylum_isolated,~ Climate_TM1*phylum*Year)
contrast(means_phylum_isolated , method="pairwise", adjust="bonferroni", by= c("phylum", "Year"))

r.squaredGLMM(mod_phylum_isolated)
```

### Connected mesocosms
```{r}
valeurs_connected <- subset(valeurs_wo2015, Connectivity_TM1 == "Connected")

valeurs_connected$Climate_TM1 <- relevel(as.factor(valeurs_connected$Climate_TM1), ref = "T38")

mod_phylum_connected <- lmer(value ~ Climate_TM1*phylum*Year + Age_class + Sex +  SVL + (1|Code) + (1|Mesocosm_TM1) , data = valeurs_connected)

Anova(mod_phylum_connected)
anova(mod_phylum_connected)

means_phylum_connected <- emmeans(mod_phylum_connected,~ Climate_TM1*phylum*Year)
contrast(means_phylum_connected, method="pairwise", adjust="bonferroni", by= c("phylum", "Year"))

r.squaredGLMM(mod_phylum_connected)
```

