---
title: "Gut Diversity - Dispersal"
author: "XXX"
date: "2022-11-17"
output:
  pdf_document:
    toc: yes
    df_print : kable
  html_document:
    number_sections: yes
    toc: yes
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE, 
                      fig.align = "center", 
                      fig.width = 8,
                      fig.height = 6, 
                      message = FALSE, 
                      warning = FALSE)
```

# Libraries
```{r libraries}
library(metabaR)
library(ggplot2)
library(forcats)
library(lme4)
library(lmerTest)
```

# Read the data and load functions
```{r, echo=TRUE}
setwd("~/data")

all_data <- readRDS("XXXMS1_cleandata.rds")

data_lizards <- all_data$samples

setwd("~/R")
source('0_XXX_sources_functions.R', chdir = TRUE)
```

# Data 

To study dispersal, we only worked on lizard from connected mesocosms. Dispersal mostly occurs during the first year of life (Le Galliard et al. 2003), but juveniles cannot be sampled for their gut microbiota at birth because of their small size. To include juveniles in the analysis, we studied the relationship between the dispersal status from year t to year t+1 and gut microbiota diversity at year t+1.

```{r}
data_lizards_connected <- subset(data_lizards, Connectivity_TM1 == "Connected")
data_lizards_connected <- subset(data_lizards_connected, Year != "2015")
```

# Plot 
<!--new variable to assign the right colors --> 
```{r, echo=FALSE}
data_lizards_connected$Clim_disp1 <- ifelse(data_lizards_connected$Climate_TM1 == "T28" & data_lizards_connected$Annual_dispersal_TM1 == "0", 
                                                 "Present_Resident", 
                                                 NA)

data_lizards_connected$Clim_disp2 <- ifelse(data_lizards_connected$Climate_TM1 == "T28" & data_lizards_connected$Annual_dispersal_TM1 == "1", 
                                                 "Present_Dispersers", 
                                                 data_lizards_connected$Clim_disp1)

data_lizards_connected$Clim_disp3 <- ifelse(data_lizards_connected$Climate_TM1 == "T38" & data_lizards_connected$Annual_dispersal_TM1 == "0", 
                                                 "Warm_Resident", 
                                                 data_lizards_connected$Clim_disp2)

data_lizards_connected$Clim_disp <- ifelse(data_lizards_connected$Climate_TM1 == "T38" & data_lizards_connected$Annual_dispersal_TM1 == "1", 
                                                 "Warm_Dispersers", 
                                                 data_lizards_connected$Clim_disp3)


data_lizards_connected$Clim_disp <- fct_relevel(data_lizards_connected$Clim_disp, c("Present_Resident", "Present_Dispersers", "Warm_Resident", "Warm_Dispersers"))

levels(data_lizards_connected$Clim_disp)
```

```{r}
Plot_Dispersal <- ggplot(data_lizards_connected, aes(x=Climate_TM1, y=q1, fill=Clim_disp)) + 
  geom_boxplot(outlier.shape = NA, alpha=0.5) +
  geom_point(aes(color=Clim_disp), 
             position = position_jitterdodge(), 
             alpha=0.35) + 
  labs(x="Climate before dispersal decision", y="exp(H)") +
  scale_fill_manual(values = c("#006699", "#990066", "#CC0000", "#990066")) +
  scale_color_manual(values = c("#006699", "#990066", "#CC0000", "#990066")) + 
  scale_x_discrete(labels = c("Present-day", "Warm")) +
  theme_fig

Plot_Dispersal
```

<!-- Save plot --> 
```{r, eval=FALSE}
setwd("~/Plots")
ggsave("Plot_Dispersal.pdf", Plot_Dispersal, width = 8, height = 5)
```
# Model 
<!--Centered years --> 
```{r}
data_lizards_connected$year2 <- ifelse(data_lizards_connected$Year =="2017", 
                            1, 
                            0)


data_lizards_connected$year3 <- ifelse(data_lizards_connected$Year =="2018", 
                            1, 
                            0)
# apply it
data_lizards_connected$year2c <- data_lizards_connected$year2 - mean(data_lizards_connected$year2)
data_lizards_connected$year3c <- data_lizards_connected$year3 - mean(data_lizards_connected$year3)
```

<!--Run model --> 
```{r}
mod_dispersal <- lmer(q1 ~ Annual_dispersal_TM1*year2c + Annual_dispersal_TM1*year3c + Annual_dispersal_TM1*Climate_TM1 + Climate_TM1*year2c + Climate_TM1*year3c + Age_class + Sex + SVL + (1|Mesocosm_TM1) + (1|Code), data = data_lizards_connected) 

dredge_function(mod_dispersal)
relative_importance(mod_dispersal)

r.squaredGLMM(mod_dispersal)
```

